\TODO{What's the title?}

### 2018-11-20 doc/Paper.tex
Update PDF to 4b4c412b5d229a27577dcdad334bdfa8c74a5e3b

@@

  Suppose that we would like to update a list at a specific position:
\begin{tabbing}\texfamily
~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Int}~\WithMath{->}{\rightarrow}~[a]~\WithMath{->}{\rightarrow}~[a]
\\
\texfamily ~~modify~f~i~xs~=~helper~[]~xs~0
\\
\texfamily ~~~{\bfseries where}~helper~before~(x~:~after)~!j
\\
\texfamily ~~~~~~~~~~~|~j~\WithMath{==}{\equiv}~i~~~~=~before~\WithMath{++}{+\!\!+}~f~x~\WithMath{++}{+\!\!+}~after
\\
\texfamily ~~~~~~~~~~~|~\textit{otherwise}~=~helper~(before~\WithMath{++}{+\!\!+}~[x])~after~(j~+~1)
\\
\texfamily ~~~~~~~~~helper~\char95 ~[]~\char95 ~=~\textit{error}~\char34 Index~out~of~bounds.\char34 
\end{tabbing}
  Here \text{\ttfamily modify} takes an update action \text{\ttfamily f}\footnote{\text{\ttfamily f} returns a list rather
  than a single element to prevent curious readers from suggesting to use a
  boxed array instead of a list.}, an index \text{\ttfamily i}, and a list \text{\ttfamily xs} and returns a
  new list with the \text{\ttfamily i}'th element replaced with the result of \text{\ttfamily f}.%
  This function "unpacks" a list, modifies one element, and "packs" the result
  into a list. If we do a lot of updates, we end up unpacking and packing the
  list over and over again -- very time-consuming for long lists. Explicitly
  working with the unpacked representation is bug-prone. A list zipper
  simplifies this.

  A zipper consists of a focus (alternatively called a hole) and surrounding
  context:
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Context}~[a]~[a]
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\end{tabbing}
  where the \text{\ttfamily ListContext} keeps track of elements to the left and to the right
  of the focus. We can now define movements:
\begin{tabbing}\texfamily
~~left~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~left~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~[]~\char95 ))~~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)
\\
\texfamily 
\\
\texfamily ~~right~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~right~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~\char95 ~[]))~~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
\end{tabbing}
  and functions for entering and leaving the zipper:
\begin{tabbing}\texfamily
~~lzEnter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~lzEnter~[]~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~lzEnter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)
\\
\texfamily 
\\
\texfamily ~~lzLeave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]
\\
\texfamily ~~lzLeave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
\end{tabbing}

  Finally, we define a local version of our \text{\ttfamily modify} function (\TODO{Boy, is
  this function ugly...})
\begin{tabbing}\texfamily
~~lzModify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~lzModify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}
\\
\texfamily ~~~~(x~:~xs)~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~ls~(xs~\WithMath{++}{+\!\!+}~rs))
\\
\texfamily ~~~~[]~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~rs~{\bfseries of}
\\
\texfamily ~~~~~~(r~:~rs')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Context}~ls~rs')
\\
\texfamily ~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~ls~{\bfseries of}
\\
\texfamily ~~~~~~~~(l~:~ls')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~l~({\itshape Context}~ls'~rs)
\\
\texfamily ~~~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Nothing}
\end{tabbing}

  \newpage
  using which we can perform multiple updates efficiently and with minimal code
  bloat\footnote{%
  Operator \text{\ttfamily \char62{}\char61{}\char62{}} comes from \text{\ttfamily Control\char46{}Monad} module in \text{\ttfamily base}
  and has the following signature:
\begin{tabbing}\texfamily
~~(>=>)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
\end{tabbing}
  }:
\begin{tabbing}\texfamily
~~modifyExample~::~{\itshape IO}~()
\\
\texfamily ~~modifyExample~=~print~\$
\\
\texfamily ~~~~lzEnter~@{\itshape Int}~>=>~right
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~right
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(const~[])
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~(+1))
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~left
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~negate)
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~return~\WithMath{.}{\circ}~lzLeave~\$
\\
\texfamily ~~~~~~[1,~2,~3,~4,~5]
\end{tabbing}

  Consider now a binary tree data structure:
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape Tree}~a
\\
\texfamily ~~~~=~{\itshape Fork}~({\itshape Tree}~a)~({\itshape Tree}~a)
\\
\texfamily ~~~~|~{\itshape Leaf}~!a
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\end{tabbing}
  A binary tree zipper is slightly more insteresting than the list zipper,
  because we can move up and down the tree as well. The zipper again consists of
  a hole (a subtree we are focused on) and its surrounding context (a path from
  the hole to the root of the tree):
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape Zipper}~a~=~{\itshape Zipper}~\char123 ~\char95 hole~::~({\itshape Tree}~a),~\char95 cxt~::~!({\itshape Context}~a)~\char125 
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\\
\texfamily 
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~a
\\
\texfamily ~~~~=~{\itshape Top}
\\
\texfamily ~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)
\\
\texfamily ~~~~|~{\itshape TreeRight}~({\itshape Tree}~a)~!({\itshape Context}~a)
\\
\texfamily ~~~~{\bfseries deriving}~({\itshape Show},~{\itshape Eq})
\end{tabbing}
  To move the zipper down, we "unpack" the current hole:
\begin{tabbing}\texfamily
~~down~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~\char95 )~=~{\itshape Nothing}
\\
\texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape TreeRight}~l~cxt)
\end{tabbing}
  \text{\ttfamily TreeContext} stores everything we need to reconstruct the hole, and \text{\ttfamily tzUp} does
  exactly that:
\begin{tabbing}\texfamily
~~up~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~up~({\itshape Zipper}~\char95 ~{\itshape Top})~=~{\itshape Nothing}
\\
\texfamily ~~up~({\itshape Zipper}~l~({\itshape Left}~cxt~r))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
\\
\texfamily ~~up~({\itshape Zipper}~r~({\itshape TreeRight}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
\end{tabbing}
  Implementation of \text{\ttfamily tzLeft}, \text{\ttfamily tzRight}, and \text{\ttfamily tzEnter} is very similar to the
  List zipper case and is left as an exercise for the reader. \text{\ttfamily tzLeave} differs
  slightly in that we now move all the way up rather than left:
\begin{tabbing}\texfamily
~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Tree}~a
\\
\texfamily ~~leave~z~=~{\bfseries case}~up~z~{\bfseries of}
\\
\texfamily ~~~~{\itshape Just}~z'~\WithMath{->}{\rightarrow}~leave~z'
\\
\texfamily ~~~~{\itshape Nothing}~\WithMath{->}{\rightarrow}~\char95 hole~z
\end{tabbing}

  The list and binary tree zipper we have considered here are homogeneous
  zippers: the type of focus does not change upon zipper movement. Such zippers
  can be a very useful abstraction. For example, a well-known window manager
  XMonad\cite{xmonad} uses a rose tree zipper to track the window under focus.
  For other tasks, however, one might need to traverse heterogeneous structures.
  A zipper that can accomodate such needs is usually called a generic zipper as
  it relies only on the generic structure of Algebraic Data Types (ADTs). One
  can view an ADT as an Abstract Syntax Tree (AST) where each node is a Haskell
  constructor rather than a syntax construct.

  The generic zipper we will use is very similar to the one presented
  in\cite{adams2010syz}. The most common technique in Haskell for supporting
  heterogeneous types is Existential Quantification. However, not every type can
  act as a hole. To support moving down the tree, we need the hole to be
  \textit{dissectible}, i.e. we would like to be able to dissect the value into
  the constructor and its arguments. Even though \text{\ttfamily Data\char46{}Data\char46{}gfold} allows us to achieve
  this, we define out own typeclass which additionally allows us to propagate
  down arbitrary constraints:
\begin{tabbing}\texfamily
~~{\bfseries class}~{\itshape Dissectible}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(a~::~{\itshape Type})~{\bfseries where}
\\
\texfamily ~~~~dissect~::~a~\WithMath{->}{\rightarrow}~{\itshape Left}~c~a
\\
\texfamily 
\\
\texfamily ~~{\bfseries data}~{\itshape Left}~c~expects~{\bfseries where}
\\
\texfamily ~~~~{\itshape LOne}~~::~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~b
\\
\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
\end{tabbing}
  For example, here is how we can make \text{\ttfamily Tree} an instance of \text{\ttfamily Disssectible}
\begin{tabbing}\texfamily
~~{\bfseries instance}~c~({\itshape Tree}~a)~=>~{\itshape Dissectible}~c~({\itshape Tree}~a)~{\bfseries where}
\\
\texfamily ~~~~dissect~({\itshape Fork}~l~r)~=~{\itshape LOne}~{\itshape Fork}~`{\itshape LCons}`~l~`{\itshape LCons}`~r
\\
\texfamily ~~~~dissect~x~=~{\itshape LOne}~x
\end{tabbing}
  We can unpack a \text{\ttfamily Fork} and the zipper will thus be able to go down. \text{\ttfamily Leaf}s,
  however, are left untouched and trying to go down from a \text{\ttfamily Leaf} will return
  \text{\ttfamily Nothing}.

>>>>

\newcommand{\ttsub}[2]{{\text{#1}_\text{#2}}}

@@

  \paragraph{Lists} Suppose that we would like to update a list at a specific
  position:
\begin{tabbing}\texfamily
~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Int}~\WithMath{->}{\rightarrow}~[a]~\WithMath{->}{\rightarrow}~[a]
\\
\texfamily ~~modify~f~i~xs~=~helper~[]~xs~0
\\
\texfamily ~~~{\bfseries where}~helper~before~(x~:~after)~!j
\\
\texfamily ~~~~~~~~~~~|~j~\WithMath{==}{\equiv}~i~~~~=~before~\WithMath{++}{+\!\!+}~f~x~\WithMath{++}{+\!\!+}~after
\\
\texfamily ~~~~~~~~~~~|~\textit{otherwise}~=~helper~(before~\WithMath{++}{+\!\!+}~[x])~after~(j~+~1)
\\
\texfamily ~~~~~~~~~helper~\char95 ~[]~\char95 ~=~\textit{error}~\char34 Index~out~of~bounds.\char34 
\end{tabbing}
  Here \text{\texfamily modify} takes an update action \text{\texfamily f}\footnote{\text{\texfamily f} returns a list rather
  than a single element to prevent curious readers from suggesting to use a
  boxed array instead of a list.}, an index \text{\texfamily i}, and a list \text{\texfamily xs} and returns a
  new list with the \text{\texfamily i}'th element replaced with the result of \text{\texfamily f}.%
  First, we ``unpack'' the list into $\text{\texfamily before\Sp =\Sp [}\ttsub{\text{\texfamily xs}}{\text{\texfamily 0}},\dots
  \ttsub{\text{\texfamily xs}}{\text{\texfamily i-1}}\text{\texfamily ]}$, $\text{\texfamily x\Sp =\Sp }\ttsub{\text{\texfamily xs}}{\text{\texfamily i}}$, and
  $\text{\texfamily after\Sp =\Sp [}\ttsub{\text{\texfamily xs}}{\text{\texfamily i+1}},\dots\text{\texfamily ]}$. Then replace \text{\texfamily x} by \text{\texfamily f(x)} and
  finally ``pack'' the result back into a single list. If we do a lot of
  updates, we end up unpacking and packing the list over and over again -- very
  time-consuming for long lists. We would thus like to benefit from fusion
  without explicitly working with the unpacked representation as it is
  bug-prone. A list zipper provides a way to achieve this.

  A zipper consists of a focus (alternatively called a hole) and surrounding
  context\footnote{`\text{\texfamily !}' in front of the type of \text{\texfamily \char95 cxt} is a
  strictness annotation enabled by the \text{\ttfamily BangPatterns} extension. Context can be
  seen as a path from the to the current focus and is always finite, i.e. there
  is no reason for it to be lazy.}:
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Context}~[a]~[a]
\end{tabbing}
  where the \text{\texfamily {\itshape Context}} keeps track of elements to the left (\text{\texfamily before} in
  \text{\texfamily modify}) and to the right (\text{\texfamily after} in \text{\texfamily modify}) of the focus. We can now
  define movements:
\begin{tabbing}\texfamily
~~left~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~left~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~[]~\char95 ))~~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~left~({\itshape Zipper}~hole~({\itshape Context}~(l~:~ls)~rs))~=~{\itshape Just}~\$
\\
\texfamily ~~~~{\itshape Zipper}~l~\$~{\itshape Context}~ls~(hole~:~rs)
\\
\texfamily 
\\
\texfamily ~~right~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~right~({\itshape Zipper}~\char95 ~~~~({\itshape Context}~\char95 ~[]))~~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~right~({\itshape Zipper}~hole~({\itshape Context}~ls~(r~:~rs)))~=~{\itshape Just}~\$
\\
\texfamily ~~~~{\itshape Zipper}~r~\$~{\itshape Context}~(hole~:~ls)~rs
\end{tabbing}
  and functions for entering and leaving the zipper:
\begin{tabbing}\texfamily
~~enter~::~[a]~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~enter~[]~~~~~~~=~{\itshape Nothing}
\\
\texfamily ~~enter~(x~:~xs)~=~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~[]~xs)
\\
\texfamily 
\\
\texfamily ~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~[a]
\\
\texfamily ~~leave~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~reverse~ls~\WithMath{++}{+\!\!+}~hole~:~rs
\end{tabbing}

  Finally, we define a local version of our \text{\texfamily modify} function (\TODO{Boy, is
  this function ugly...})
\begin{tabbing}\texfamily
~~modify~::~(a~\WithMath{->}{\rightarrow}~[a])~\WithMath{->}{\rightarrow}~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~modify~f~({\itshape Zipper}~hole~({\itshape Context}~ls~rs))~=~{\bfseries case}~f~hole~{\bfseries of}
\\
\texfamily ~~~~(x~:~xs)~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~x~({\itshape Context}~ls~(xs~\WithMath{++}{+\!\!+}~rs))
\\
\texfamily ~~~~[]~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~rs~{\bfseries of}
\\
\texfamily ~~~~~~(r~:~rs')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Context}~ls~rs')
\\
\texfamily ~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\bfseries case}~ls~{\bfseries of}
\\
\texfamily ~~~~~~~~(l~:~ls')~\WithMath{->}{\rightarrow}~{\itshape Just}~\$~{\itshape Zipper}~l~({\itshape Context}~ls'~rs)
\\
\texfamily ~~~~~~~~[]~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Nothing}
\end{tabbing}
  using which we can perform multiple updates efficiently and with minimal code
  bloat\footnote{%
  Operator \text{\ttfamily \char62{}\char61{}\char62{}} comes from \text{\texfamily {\itshape {\itshape Control}.Monad}} module in \text{\ttfamily base}
  and has the following signature:%
\begin{tabbing}\texfamily
~~(>=>)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
\end{tabbing}
%
  }:
\begin{tabbing}\texfamily
~~modifyExample~::~{\itshape IO}~()
\\
\texfamily ~~modifyExample~=~print~\$
\\
\texfamily ~~~~enter~@{\itshape Int}~>=>~right
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~right
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(const~[])
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~(+1))
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~left
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~modify~(return~\WithMath{.}{\circ}~negate)
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~return~\WithMath{.}{\circ}~leave~\$~[1,~2,~3,~4,~5]
\end{tabbing}

  \paragraph{Trees} Consider now a binary tree data structure:
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape Tree}~a~=~{\itshape Leaf}~!a~|~{\itshape Fork}~({\itshape Tree}~a)~({\itshape Tree}~a)
\end{tabbing}
  A binary tree zipper is slightly more insteresting than the list zipper,
  because we can move up and down the tree as well as left and right. The zipper
  again consists of a hole (a subtree we are focused on) and its surrounding
  context (a path from the hole to the root of the tree):
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape Zipper}~a~~=~{\itshape Zipper}~\char123 ~\char95 hole~::~{\itshape Tree}~a,~\char95 cxt~::~!({\itshape Context}~a)~\char125 
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~a~=~{\itshape Top}
\\
\texfamily ~~~~~~~~~~~~~~~~~|~{\itshape Left}~!({\itshape Context}~a)~({\itshape Tree}~a)
\\
\texfamily ~~~~~~~~~~~~~~~~~|~{\itshape Right}~({\itshape Tree}~a)~!({\itshape Context}~a)
\end{tabbing}
  To move the zipper down, we ``unpack'' the current hole:
\begin{tabbing}\texfamily
~~down~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~down~({\itshape Zipper}~({\itshape Leaf}~\char95 )~~~\char95 )~~~=~{\itshape Nothing}
\\
\texfamily ~~down~({\itshape Zipper}~({\itshape Fork}~l~r)~cxt)~=~{\itshape Just}~\$~{\itshape Zipper}~r~({\itshape Right}~l~cxt)
\end{tabbing}
  \text{\texfamily {\itshape Context}} stores everything we need to reconstruct the hole, and \text{\texfamily up} does
  exactly that:
\begin{tabbing}\texfamily
~~up~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Maybe}~({\itshape Zipper}~a)
\\
\texfamily ~~up~({\itshape Zipper}~\char95 ~{\itshape Top})~=~{\itshape Nothing}
\\
\texfamily ~~up~({\itshape Zipper}~l~({\itshape Left}~cxt~r))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
\\
\texfamily ~~up~({\itshape Zipper}~r~({\itshape Right}~l~cxt))~=~{\itshape Just}~\$~{\itshape Zipper}~({\itshape Fork}~l~r)~cxt
\end{tabbing}
  Implementations of \text{\texfamily left}, \text{\texfamily right}, and \text{\texfamily enter} are very similar to the
  list zipper case and are left as an exercise for the reader. \text{\texfamily leave} differs
  slightly in that we now move all the way up rather than left:
\begin{tabbing}\texfamily
~~leave~::~{\itshape Zipper}~a~\WithMath{->}{\rightarrow}~{\itshape Tree}~a
\\
\texfamily ~~leave~z~=~{\bfseries case}~up~z~{\bfseries of}
\\
\texfamily ~~~~{\itshape Just}~z'~\WithMath{->}{\rightarrow}~leave~z'
\\
\texfamily ~~~~{\itshape Nothing}~\WithMath{->}{\rightarrow}~\char95 hole~z
\end{tabbing}

  \paragraph{Generic Zipper} The list and binary tree zippers we have considered
  so far are homogeneous zippers: the type of the focus does not change upon zipper
  movement. Such a zipper can be a very useful abstraction. For example, a
  well-known window manager XMonad\cite{xmonad} uses a rose tree zipper to track
  the window under focus. For other tasks, however, one might need to traverse
  heterogeneous structures. A zipper that can accomodate such needs is usually
  called a \textit{generic zipper} as it relies only on the generic structure of
  Algebraic Data Types (ADTs). One can view an ADT as an Abstract Syntax Tree
  (AST) where each node is a Haskell constructor rather than a syntax construct.

  The generic zipper we will use is very similar to the one presented
  in\cite{adams2010syz}. The most common technique in Haskell for supporting
  heterogeneous types is Existential Quantification. However, not every type can
  act as a hole. To support moving down the tree, we need the hole to be
  \textit{dissectible}, i.e. we would like to be able to dissect the value into
  the constructor and its arguments. Even though \text{\texfamily {\itshape Data}.{\itshape Data}.gfoldl} allows us to
  achieve this, we define out own typeclass which additionally allows us to
  propagate down arbitrary constraints:
\begin{tabbing}\texfamily
~~{\bfseries class}~{\itshape Dissectible}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(a~::~{\itshape Type})~{\bfseries where}
\\
\texfamily ~~~~dissect~::~a~\WithMath{->}{\rightarrow}~{\itshape Left}~c~a
\\
\texfamily 
\\
\texfamily ~~{\bfseries data}~{\itshape Left}~c~expects~{\bfseries where}
\\
\texfamily ~~~~{\itshape LOne}~~::~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~b
\\
\texfamily ~~~~{\itshape LCons}~::~(c~b,~{\itshape Dissectible}~c~b)
\\
\texfamily ~~~~~~~~~~=>~{\itshape Left}~c~(b~\WithMath{->}{\rightarrow}~expects)~\WithMath{->}{\rightarrow}~b~\WithMath{->}{\rightarrow}~{\itshape Left}~c~expects
\end{tabbing}
  For example, here is how we can make \text{\texfamily {\itshape Tree}} an instance of \text{\texfamily {\itshape Disssectible}}
\begin{tabbing}\texfamily
~~{\bfseries instance}~c~({\itshape Tree}~a)~=>~{\itshape Dissectible}~c~({\itshape Tree}~a)~{\bfseries where}
\\
\texfamily ~~~~dissect~({\itshape Fork}~l~r)~=~{\itshape LOne}~{\itshape Fork}~`{\itshape LCons}`~l~`{\itshape LCons}`~r
\\
\texfamily ~~~~dissect~x~=~{\itshape LOne}~x
\end{tabbing}
  We can unpack a \text{\texfamily {\itshape Fork}} and the zipper will thus be able to go down. \text{\texfamily {\itshape Leaf}}s,
  however, are left untouched and trying to go down from a \text{\texfamily {\itshape Leaf}} will return
  \text{\texfamily {\itshape Nothing}}.

  To allow the zipper to move left and right, we need a means to encode
  arguments to the right of the hole. Following Adams et al, we define a GADT
  representing constructor arguments to the right of the hole:
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape Right}~c~provides~r~{\bfseries where}
\\
\texfamily ~~~~{\itshape RNil}~~::~{\itshape Right}~c~r~r
\\
\texfamily ~~~~{\itshape RCons}~::~(c~b,~{\itshape Dissectible}~c~b)
\\
\texfamily ~~~~~~~~~~=>~b~\WithMath{->}{\rightarrow}~{\itshape Right}~c~provides~r~\WithMath{->}{\rightarrow}~{\itshape Right}~c~(b~\WithMath{->}{\rightarrow}~provides)~r
\end{tabbing}
  For example, for a tuple \text{\texfamily ({\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int},\Sp {\itshape Int})}, we can have
\begin{tabbing}\texfamily
~~lefts~=~{\itshape LOne}~(,,,,,)~`{\itshape LCons}`~1~`{\itshape LCons}`~2~`{\itshape LCons}`~3
\\
\texfamily ~~hole~=~4
\\
\texfamily ~~rights~=~5~`{\itshape RCons}`~6~`{\itshape RCons}`~{\itshape RNil}
\end{tabbing}
  generalising this a little, we arrive at:
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape LocalContext}~c~hole~rights~parent~=
\\
\texfamily ~~~~{\itshape LocalContext}~!({\itshape Left}~c~(hole~\WithMath{->}{\rightarrow}~rights))~!({\itshape Right}~c~rights~parent)
\\
\texfamily 
\\
\texfamily ~~{\bfseries data}~{\itshape Context}~::~({\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Type}~{\bfseries where}
\\
\texfamily ~~~~{\itshape RootContext}~::~\char'24~c~root.~{\itshape Context}~c~root~root
\\
\texfamily ~~~~(:>)~::~\char'24~c~parent~root~hole~rights.~(c~parent,~{\itshape Dissectible}~c~parent)
\\
\texfamily ~~~~~~~~~=>~!({\itshape Context}~c~parent~root)
\\
\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~!({\itshape LocalContext}~c~hole~rights~parent)
\\
\texfamily ~~~~~~~~~\WithMath{->}{\rightarrow}~{\itshape Context}~c~hole~root
\end{tabbing}
  And just like before the \text{\texfamily {\itshape Zipper}} is a product of the hole and context:
\begin{tabbing}\texfamily
~~{\bfseries data}~{\itshape Zipper}~(c~::~{\itshape Type}~\WithMath{->}{\rightarrow}~{\itshape Constraint})~(root~::~{\itshape Type})~=
\\
\texfamily ~~~~\char'24~hole.~(c~hole,~{\itshape Dissectible}~c~hole)~=>
\\
\texfamily ~~~~~~{\itshape Zipper}~\char123 ~\char95 zHole~::~!hole
\\
\texfamily ~~~~~~~~~~~~~,~\char95 zCxt~~::~!({\itshape Context}~c~hole~root)
\\
\texfamily ~~~~~~~~~~~~~\char125 
\end{tabbing}

  Implementation of movements is quite straightforward and is left out. Please,
  refer to (\TODO{github repo}) for complete code. We now consider a rather
  interesting application of generic zipper: embedding of attribute grammars.
